The binary liquid model for gravity-driven microchannel simulations

### 2010-10-25 Text/paper.tex
Added the capillary number text. Next iteration is the viscous capillary number simulations.

\subsection{The influence of the wall gradient}
The results should not depend on the wall wettability. Wettability is defined
through the phase gradient \cite{pooley-contact}, $\partial_n \phi$. This
section is to examine the wall gradient influence on simulations. We take large
enough resolution for simulations to be consistent. Simulation domain was
chosen as $177\mathrm{x}2626$ and the initial film width as
$12$ lattice boltzmann units, which correspond to the predicted film thickness.
We examined $11$ different values for the wall
gradient ranging from $-1$ to $1$. The results are consistent among all the
wall gradients given that the interface is resolved in the right way. For the
given wall gradients the values of interface thicknesses are
$0.0633$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0633$,$0.0632$,
$0.0631$. The results are different in 4th digit. For the wall gradients $0.8$
and $1.0$ the
simulations are unstable. One can see the simulated profiles in Fig.
\ref{fig:gradients:profiles} The negative values of the phase gradient are
preferable, as
far as the phase of the liquid adjacent to the wall has the value of $1$. That
means that the phase values near the wall are above $1$ and do not interact
with the bubble which has phase value $-1$.  In case of positive
gradients the values near the wall are below $1$ and the gradient profile can
fuse with the values of the gas phase as $-1$, creating the wall stickyness,
unphysical one.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Wall/phase_grad_profiles.eps}
\caption{Phase profiles for different wall gradients. One can see if the width
is properly resolved then the wettability of the wall doesn't play any role
affecting the film thickness.
\label{fig:gradients:profiles}
}
\end{figure}
\section{The influence of different viscosities}
The results obtained earlier are taken for the liquiqs viscosity ratio $10$.
We performed the same simulations for the liquids viscosity ratio as $20$.
Parameters were taken as $\tau_{liquid}=4.5$ and $\tau_{bubble}=0.7$. 
{\color{red} The results showed that blah-blah-blah - I guess will be the same}.

\section{Capillary number region}
The purpose of this section is to validate the correlations of
\citet{giavedoni-numerical}, Fig. \ref{fig:giavedoni:planar}, and
\citet{heil-bretherton}, Fig. \ref{fig:heil:planar}, for the range of capillary
numbers. Because of limiting computational factors as memory and the
performance we cannot go to the range of small
capillary numbers. However, all the other methods were validated across the
small capilary numbers and extended for the range of capillary numbers from
$0.005$ to $1.0$. Our goal is to compare the film thickness over the capillary
range from $0.03$ to $1.0$.  For the
benchmark
results we
suggest the following Capillary numbers as:
\begin{equation}
0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8,1.0\, .
\end{equation}
The corresponding approximate film widths from \cite{giavedoni-numerical} are as
follows:
\begin{equation}
0.04,0.06,0.08,0.1,0.12,0.13,0.15\,.
\end{equation}
To perform numerical simulations we refer
to
grid refinement section where we obtained that the interface width should be
at most $50$ percent of the interface width. Therefore the grid was chosen as
$202 \mathrm{x} 3001$. $5$ lattice boltzmann units do not occupy more than $60$
percent of the effective width. Because the simulation get unstable with
smaller grids and larger gradients, all the capillary number simulations were
performed on the grid chosen. To properly initialize the force gradient the
proportionality law was utilized. For instance, with the forcing
$6\,10^{-6}/16$ was chosen for the grid to obtain predicted capillary number as
$0.005$. Therefore, one can obtain the initial forcing for other grids using
simple correlations: 
\begin{equation}
\begin{aligned}
&Ca_{lit} \propto U_{bubble}\\
&U_{bubble} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\ 
&Ca_{lit} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{Ca_{lit}}{N_y^2},
\end{aligned}
\end{equation}
where the index $lit$ stands for the predicted capillary number
\cite{giavedoni-numerical,heil-bretherton}. The grid number $Ny$ is not
involved as because the simulations are conducted on the same grid.
Therefore,the pressure gradient can be obtained through the capillary numbers
ratio:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d} x}=6\,10^{-6}/16 \frac{Ca_{lit}}{0.05}
\end{equation}
The film width is initialized through the capillary numbers ratio as well:
\begin{equation*}
w=12 \frac{Ca_{lit}}{0.05}
\end{equation*}
The results were obtained after $200,000$ steps.
The physical
time is as follows:
\begin{equation}
\begin{aligned}
&\Delta t=\frac{U_{slug,LB} \Delta X}{U_{phys}} \propto \frac{1}{N_y-2}\\
&N_{iter}=200,000 \frac{N_y-2}{200}\approx 1000 N_y\\
%&N_{output}=51 Ny\\
\end{aligned}
\end{equation}

{\color{red} That's why we need better grid for larger Capillary numbers. Steady
simulations were obtained for the following Capillary numbers
$0.03,0.05,0.08,0.1$  }with the film thicknesses as
{\color{red} $0.0337433850478, 0.063192148642, 0.0799078589211,
0.0891674460655$} which are
quite close to the predicted one $0.04,0.06,0.08,0.1$. The thing is to identify
the Capillary numbers associated with those numbers. The velocity in the
centers are $0.00209258774,0.003826643811,0.006506324324,0.008333889077$. The
Capillary numbers associated with these velocities are $0.036992074529545796, 
0.067646144698590829,0.1150166512524845,0.14732373700007995$, which again a bit
overpredictive but consistent. The fig.
\ref{fig:capillary:simple} represents the thickness against the Capillary
number which mimics the curve behavior for the Capillary number dependance as
indicated in Fig. \ref{fig:giavedoni:planar}.
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/capillary_thickness_simple.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/capillary_range.eps}
\caption{{\color{red} This one to eliminate - just present here
comparison between our simulations in the longer range of capillary
numbers and Giavedoni}Thickness versus the capillary
number for different scalings based on
the average capillary number. \label{fig:capillary:simple}}
\end{figure}

\subsection{The influence of the force}

{\color{red} Michal if you find it useful - use this section. My suggestion is
to eliminate it.}
\begin{description}
 \item 
On the grid $198\mathrm{x}2941$ and the initial film width as $20$ units
the
influence of the force is studied. $16$ different values of force are examined,
which have the following expression $\frac{0.000006}{16} i$, where $i=1 \dots
16$. The system is allowed to come to the equilibrium after $200,000$ steps. The
goal is to study the influence of the force on the velocity and therefore the
capillary number. Unfortunately in the first case only a few simulations are
stable, i.e.
$i=1,2$ is stable. The results for the interface thicknesses are
$0.0700122222668,0.0985152303286$ with the velocities associated as
$0.004490013864,0.009597299475$. Those numbers correspond to the bubble
capillary numbers as $0.079372981271400328, 0.16965788849626481$. The average
capillary numbers are $0.07381636,0.14095644$.

\item
On the grid $394\mathrm{x}5881$ and the initial film width as $20+4 i$,where
$i=1 \dots 5$ units the
influence of the force is studied. The force is scheduled as
$\frac{0.000006}{128}+i \frac{0.000006}{640}$.The intention is to cover the
calculation of the Capillary numbers in the range $0.03,0.035,0.04,0.045,0.05$. 
$5$ different values of force are examined. The system is allowed to come to the
equilibrium after $400,000-40,000 i$ steps. The associated thicknesses were as
it follows
$0.0488915027016,0.0538098132568,0.0587100833678,0.0615470184196,
0.0661331899123$. Bubble velocities are
$0.002592856205,0.003077834398,0.003562975124,0.004034533823,0.004470657603$.
Bubble Capillary numbers are $0.04583566,  0.05440894,  0.0629851 , 
0.07132116, 0.07903081$.
\end{description}

>>>>

\subsection{The influence of the wall gradient}
The results should not depend on the wall wettability. Wettability is defined
through the phase gradient \cite{pooley-contact}, $\partial_n \phi$. This
section is to examine the wall gradient influence on simulations. We take large
enough resolution for simulations to be consistent. Simulation domain was
chosen as $177\mathrm{x}2626$ and the initial film width as
$12$ lattice boltzmann units, which correspond to the predicted film thickness.
We examined $11$ different values for the wall
gradient ranging from $-1$ to $1$. The results are consistent among all the
wall gradients given that the interface is resolved in the right way. For the
given wall gradients the values of interface thicknesses are
$0.0633$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0634$,$0.0633$,$0.0632$,
$0.0631$. The results are different in 4th digit. For the wall gradients $0.8$
and $1.0$ the
simulations are unstable. One can see the simulated profiles in Fig.
\ref{fig:gradients:profiles} The negative values of the phase gradient are
preferable, as
far as the phase of the liquid adjacent to the wall has the value of $1$. That
means that the phase values near the wall are above $1$ and do not interact
with the bubble which has phase value $-1$.  In case of positive
gradients the values near the wall are below $1$ and the gradient profile can
fuse with the values of the gas phase as $-1$, creating the wall stickyness,
unphysical one.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Wall/phase_grad_profiles.eps}
\caption{Phase profiles for different wall gradients. One can see if the width
is properly resolved then the wettability of the wall doesn't play any role
affecting the film thickness.
\label{fig:gradients:profiles}
}
\end{figure}

\section{Capillary number region}
The purpose of this section is to validate the correlations of
\citet{giavedoni-numerical}, Fig. \ref{fig:giavedoni:planar}, and
\citet{heil-bretherton}, Fig. \ref{fig:heil:planar}, for the range of capillary
numbers. Because of limiting computational factors as memory and the
performance we cannot go to the range of small
capillary numbers. However, all the other methods were validated across the
small capilary numbers and extended for the range of capillary numbers from
$0.005$ to $10.0$. Our goal is to compare the film thickness over the capillary
range from $0.03$ to $0.8$, which is a computationally reasonable task.  For the
benchmark
results we
suggest the following Capillary numbers as:
\begin{equation}
0.03,0.05,0.08,0.1,0.2,0.4,0.6,0.8\, .
\end{equation}
The corresponding approximate film widths from \cite{giavedoni-numerical} are as
follows:
\begin{equation}
0.04,0.06,0.08,0.1,0.12,0.13,0.15\,.
\end{equation}
To perform numerical simulations we refer
to
grid refinement section where we obtained that the interface width should be
at most $50$ percent of the interface width. Therefore the grid was chosen as
$202 \mathrm{x} 3001$. $5$ lattice boltzmann units do not occupy more than $60$
percent of the effective width. Because the simulation get unstable with
smaller grids and larger gradients, all the capillary number simulations were
performed on the grid chosen. To properly initialize the force gradient the
proportionality law was utilized. For instance, with the forcing
$6\,10^{-6}/16$ was chosen for the grid to obtain predicted capillary number as
$0.005$. Therefore, one can obtain the initial forcing for other grids using
simple correlations: 
\begin{equation}
\begin{aligned}
&Ca_{lit} \propto U_{bubble}\\
&U_{bubble} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\ 
&Ca_{lit} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{Ca_{lit}}{N_y^2},
\end{aligned}
\end{equation}
where the index $lit$ stands for the predicted capillary number
\cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
involved as because the simulations are conducted on the same grid.
Therefore,the pressure gradient can be obtained through the capillary numbers
ratio:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d} x}=6\,10^{-6}/16 \frac{Ca_{lit}}{0.05}
\end{equation}
The film width is initialized through the capillary numbers ratio as well:
\begin{equation*}
w=12 \frac{Ca_{lit}}{0.05}
\end{equation*}
The results were obtained after $200,000$ steps. Film thicknesses were obtained as $0.042$,$0.066$, $0.085$, $0.077$, $0.123$, $0.153$, $0.164$, $0.173$ with the bubble center velocities as $0.0022$, $0.0041$, $0.0068$, $0.0055$, $0.0185$, $0.0388$, $0.0592$, $0.0782$. Based on the obtained center velocities the capillary numbers 
are formed as $0.039$,$0.072$,$0.120$,$0.098$,$0.328$,$0.686$,$1.046$,$1.383$. Those capillary numbers are overpredicted but the obtained lengths are overpredicted as well due to the improper body force imposement. 
We extracted data from \cite{giavedoni-numerical,heil-bretherton} and compared it with the simulation. The results are shown in Fig. \ref{fig:capillary:comparison}.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Capillary/capillaries_comparison.eps}
\caption{The comparison between simulation results and results from \cite{giavedoni-numerical} and \cite{heil-bretherton}. One can see the reasonable agreement. The film thickness against the capillary number range is presented.\label{fig:capillary:comparison}}
\end{figure}

\section{The influence of different viscosities}
The results obtained earlier are taken for the liquiqs viscosity ratio $10$.
We performed the same simulations for the liquids viscosity ratio as $20$.
Parameters were taken as $\tau_{liquid}=4.5$ and $\tau_{bubble}=0.7$. 
{\color{red} The results showed that blah-blah-blah - I guess will be the same}.
